---
title: "Project"
output:
  html_document: default
  pdf_document: default
  word_document: default
---


Analysis of Housing Prices

Questions of Interest:

Does lotsize have a significant effect on housing price. We suspect it does, and if that's true we'd like to know if any other predictors such as: the number of bathrooms, stories, and availability of air conditioning.

Regression Method:

Plot Housing prices against lotsize and use an information criterion (AIC) to help determine which other predictors are useful to our model.

```{r}

library(Ecdat)
library(MASS)
data("Housing")

df<-Housing
df$bedrooms=as.factor(df$bedrooms)
df$bathrms=as.factor(df$bathrms)
df$stories=as.factor(df$stories)
df$driveway=as.factor(df$driveway)
df$recroom=as.factor(df$recroom)
df$fullbase=as.factor(df$fullbase)
df$gashw=as.factor(df$gashw)
df$airco=as.factor(df$airco)
df$garagepl=as.factor(df$garagepl)
df$prefarea=as.factor(df$prefarea)
attach(df)


plot(lotsize,price, main= "Price vs. Lot Size")
fit<-lm(price~lotsize)

```

Plotting the Residual vs. Fitted scatter plot:


```{r }
plot(fitted(fit),resid(fit))
abline(0,0)

qqnorm(price)
qqline(price)


```
The Residual plot shows some slight fanning, indicating a break in LINE conditions. Unequal variance is likely present. The Normal Q-Q Plot is obviously non linear and breaks normality condition. We will use the box cox method to find a suitable transformation on our response.

```{r }

bc<-boxcox(fit,lambda=seq(-1,1),length=10)
estimate<-bc$x[which.max(bc$y)]
estimate

```
After applying the box cox method, we chose a log function to transform our response data. Now we plot the residuals of the transformed linear model.


```{r }
fit.trans<-lm(log(price)~lotsize)
plot(fit.trans,c(1,2))
```

The transformed residual plot has a horizontal band around the 0 line, with no discernable pattern in the plot. High predictor values trend to lower residual values, but are relatively normal. Thus we satisfy the Linearity and Equal Variane LINE conditions.

The transformed Q-Q plot is linear and satisfies our Normality LINE condition.


Now we will check the independence of our transformed data by plotting residuals vs. order.

```{r }
order<-as.factor(c(1:546))
plot(order,resid(fit.trans))
```

The residual vs. order plot is completely random, with no obvious patterns, thus we satisfy our Independence LINE condition.


Now we will conduct a t test to determine if we should reject or accept our null hypothesis. Our hypothesis is stated below.


$H_{0}$: $\beta_{1}$=0

$H_{1}$: $\beta_{1}$=/=0


```{r }
summary(fit.trans)
```

We get a student t value of 15.08, with p value < 2e-16. Which is < $\alpha$ = 0.05. This allows us to reject the null hypothesis and assume lotsize has a significant effect on our response, price.


Regression Analysis, Building Our Model:
Using a scatterplot matrix we will look for the correlation between price and lotsize. Lotsize is the only numeric predictor, which is why we run a correlation matrix on it.

```{r }
library("PerformanceAnalytics")

fit.full<-lm(price~lotsize+bedrooms+bathrms+stories+driveway+recroom+fullbase+gashw+airco+garagepl+prefarea)
pairs(price~lotsize+bedrooms+bathrms+stories)
df2<-data.frame(price,lotsize,as.numeric(bedrooms),as.numeric(bathrms),as.numeric(stories),as.numeric(driveway),as.numeric(recroom),as.numeric(fullbase),as.numeric(gashw),as.numeric(airco),as.numeric(garagepl),as.numeric(prefarea))

chart.Correlation(df2, histogram=TRUE, pch=19)


```
lotsize (0.52), bathrooms (0.52), and airconditioning (0.45) have the three highest correlation values.

Using the Akaike Information Criterion we will select the final predictors for the model. 

```{r }
#fit.red = lm(price~1)
#fit.full = lm(price~lotsize+bedrooms+bathrms+stories+driveway+recroom+fullbase+gashw+airco+garagepl+prefarea)
#step(fit.red, scope = list(lower=fit.red, upper=fit.full))

mod0<-lm(price~1)
add1(mod0,~.+lotsize+bedrooms+bathrms+stories+driveway+recroom+fullbase+gashw+airco+garagepl+prefarea,test="F")
mod1=update(mod0,~.+lotsize)
plot(bathrms,resid(mod1))
add1(mod1,~.+bedrooms+bathrms+stories+driveway+recroom+fullbase+gashw+airco+garagepl+prefarea,test="F")
mod2=update(mod1,~.+bathrms)
plot(airco,resid(mod2))
add1(mod2,~.+bedrooms+stories+driveway+recroom+fullbase+gashw+airco+garagepl+prefarea,test="F")
mod3=update(mod2,~.+airco)
plot(stories,resid(mod3))
anova(mod3)

```

The AIC method suggests we use lotsize, bathrms, and airco as the predictors in our model. The residual vs. next best predictor plots helped us determine when to add another predictor based on if a pattern is present. The last plot, stories vs. residuals of model including: lotsize, bathrms, airco has a nominal pattern which is why we stopped adding predictors.

Research Question:
Is a model containing at least one of our predictors from the dataset useful in predicting the price of a property.

Yes. To test this we conduct the following hypothesis test:

$\beta_{lotsize}$ = $\beta_{bathrms}$ = $\beta_{airco}$ = 0 vs. At least one $\beta_{k}$ $\neq$ 0

We conduct the test with a general F-test. Our full model is $Y_{i}$ = $\beta_{0}$ + $\beta_{lotsize}x_{i,lotsize}$ + $\beta_{bathrms}x_{i,bathrms}$ + $\beta_{airco}x_{i,airco}$ and the reduced model is $Y_{i}$ = $\beta_{0}$ + $\epsilon_{i}$.


```{r }
fit.red<-lm(price~1)
fit.full<-mod3
anova(fit.red,fit.full)

``` 

The F value is 130.13, with a p values < 2.2e-16, which is lower than $alpha$ = 0.05. Thus we reject the null hypothesis, and conclude at least one predictor from our full model is useful in predicting price.


Research Question:
Is a reduced version of our full model (including lotsize, bathrooms, and air conditioning) sufficient in estimating price?

Hypothesis:

$H_{1}$: The full model, price = lotsize + bathrooms + airconditioning

1. $H_{0}$: The reduced model, price = lotsize + bathrooms. 

```{r }
fit.red.x1x2<-lm(price~lotsize+bathrms)
anova(fit.red.x1x2,fit.full)

```
2. $H_{0}$: The reduced model, price = lotsize + air conditioning 
```{r }
fit.red.x1x3<-lm(price~lotsize+airco)
anova(fit.red.x1x3,fit.full)

```
3. $H_{0}$: The reduced model, price = bathrooms + air conditioning 
```{r }
fit.red.x2x3<-lm(price~bathrms+airco)
anova(fit.red.x2x3,fit.full)


```
4. $H_{0}$: The reduced model, price = lotsize 
```{r }
fit.red.x1<-lm(price~lotsize)
anova(fit.red.x1,fit.full)

```
5. $H_{0}$: The reduced model, price = bathrooms
```{r }

fit.red.x2<-lm(price~bathrms)
anova(fit.red.x2,fit.full)
```
6. $H_{0}$: The reduced model, price = air conditioning 
```{r }
fit.red.x3<-lm(price~airco)
anova(fit.red.x3,fit.full)

```

All our General F Tests reject the null hypothesis for possible reduced models, thus price = lotsize + bathrooms + air conditioning will be our final model.

Residual Analysis of the Final Model:
```{r }
plot(fit.full,c(1,2))
```
The final model doesn't meet the LINE conditions, specifically the normal Q-Q plot indicates non-normal error. We will use the box cox method to transform price.

```{r }
bc<-boxcox(fit.full,lambda=seq(-1,1),length=10)
estimate<-bc$x[which.max(bc$y)]
estimate
``` 
The box cox method suggests a $\lambda$ = 0.1111111. 

```{r }
fit.full<-lm(price^(0.1111111)~lotsize+bathrms+airco,data=Housing)
plot(fit.full,c(1,2))
```

Now we will use the model to predict the price of house with 30,000 square feet.
```{r }
library(ggplot2)
housingPrediction<-predict(fit.full,Housing)
housingdResid<-housingPrediction-df$price

ggplot(df, aes(x=seq(1, length(df$price)), y=price))+
geom_point(aes(color=abs(housingdResid), size=abs(housingdResid)))+
geom_point(aes(y=housingPrediction), shape=1)+
geom_segment(aes(xend=seq(1, length(df$price)), yend=housingPrediction), alpha=0.2)+
scale_color_continuous(low="orange", high="purple")+
labs(title="Predicted vs Observed Price", x="Index", y="Price", size="| Residuals |")+
theme_bw()+
guides(color = FALSE)

```